Uppangala tree functionnal: Article draft

### 2016-07-01 inst/doc/RLQ.Rmd
v0.1.9016 RLQ.Rmd + cosaspect fix

---
title: 'Uppangala tree functionnal traits: Preliminray Results'
author: "Sylvain Schmitt"
date: '`r Sys.Date()`'
output: pdf_document
bibliography: refs.bib
---

@@

## Communities

```{r 5 Com}
library(FD)
com <- CWT(Trees, com)
x <- data.frame(PFT_sp[-1], row.names = PFT_sp$Sp_Code)
a <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
a[is.na(a)] <- 0
x <- x[names(a),]
FD <- as.data.frame(dbFD(x, a, corr = 'cailliez', calc.CWM = F, print.pco = F, messages = F))[-c(1:2,4)]
names(FD) <- paste0(names(FD), '.fd')
FD$id <- row.names(FD)
com$abundance <- merge(com$abundance, FD)
```

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. For all those community, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES). Weighted mean and variances were realized by abundance, presence-absence, or basal area. In addition functionnal diversity indices weighted by abundances were computed with the `FD` package [@laliberte_distance-based_2010] for our communities: functional richness (FRic), functional evenness (FEve), and functional divergence (FDiv) and Rao's quadratic entropy (Q) (cite Villéger et al. (2008) et Botta-Dukát 2005).

Spatial variograms of community weighted means and variances didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$. Functionnal diversity indices show also a plateau around $70 m$.

```{r Variograms}
library(gstat)
dist <- variogram(WD.cwm~1, as(com$abundance, 'SpatialPointsDataFrame'))$dist
par(mfrow = c(2,1))
for(i in 1:2){
  traits <- paste0(c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES'), c('.cwm','.cwv')[i])
  plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', main = c('(a)', '(b)')[i], xlab = 'distance (m)', ylab = 'relative variance (%)')
  for(j in 1:length(com)){
    var <- data.frame(lapply(lapply(as.list(traits), function(x){variogram(formula(paste(x, '~ 1')), as(com[[j]], 'SpatialPointsDataFrame'))}), function(x){x[3]}))
    var <- data.frame(apply(var, 2, function(x){x/max(x)}))
    names(var) <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
    colramp <- rainbow(length(var))
    for(i in 1:length(var)){
      lines(var[,i] ~ dist, col = colramp[i], lty = j + 1)
    }
  }
legend('bottomright', names(var), fill = colramp, cex = 0.7)
legend('bottomleft', names(com), lty = c(2:4), cex = 0.7)
}
```

Figure 4: Trait variograms for all communities.*Variograms were initiated on* $20*20 m^2$ *communities for all traits with both community weighted means* **(a)** *and variances* **(b)** *(by abundance, presence-absence and basal area). Some variaograms seem to reach a plateau around* $70 m$.

In order to see if environmental biotic factors were influencing functionnal traits or if environmental biotic factors were the results of functionnal traits we used structural equation modelling on community weighted means by abundance. Three latent variables were defined: *Abiotic environment* defined with our 4 selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits. Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*). Retained model was the first one (**m1**) with a significantly lower Aikaike's Criterion Indice on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)

```{r SEM}
library(lavaan)
m1 <- '
Abiotic =~ Slope + Curvature + Wetness + SW
Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
Biotic =~ BA + Canopy
Traits ~ Abiotic + Biotic
'
m2 <- '
Abiotic =~ Slope + Curvature + Wetness + SW
Traits =~ WD.cwm + Thick.cwm + LA.cwm + LDMC.cwm + SLA.cwm
Biotic =~ BA + Canopy
Biotic ~ Traits + Abiotic
'
fit1 <- sem(m1, data = com$abundance@data)
fit2 <- sem(m2, data = com$abundance@data)
semPlot::semPaths(fit1,
                  what = 'col',
                  whatLabels = 'std',
                  style = 'lisrel',
                  residuals = T,
                  nCharNodes = 0,
                  nCharEdges = 0,
                  groups='manlat',
                  pastel = T,
                  ask = F)
```

Figure 5: Path diagram for structural equation modelling. *Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic. Arrow values are indicating standardized parameter estimates*.

Links between both community weighted means and variances with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables. Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables (slope, curvature, wetness and southwesterness for abiotic factors, and canopy height and basal area for biotic one). A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix). Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance. Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 4 and 5 (respectively for community weighted means and variances). Same procedure was also applied to functionnal diversity indices (FD) from our ocmmunities, results are presented in table 6.

An assessment of spatial autocorrelation (SAC) of the community weighted means and variances and functionnal diversity indices was conducted using Moran's I statistic. Only half of community weighted means and variances were significantly positively spatially autocorrelated (see table 3) with low values (Moran's I from 0.01 to 0.04). Leaf economic spectrum and dry matter content weighted means as wood density weighted variances are slightly spatially correlated suggesting a spatial filter due to environment or local history pressure (as dispersion limitation). On the opposite, wood density and thickness community weighted means as wood economic spectrum community weighted variances possess a stochastic spatial distribution implying that they're not shaped by factors with a spatial organization. Finally remaining traits community weighted means and variances are spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter. As a conclusion both community weighted means and variances are most of the time stochastic or driven by dispersion. Moreover all functionnal diversity indices (FD) are spatially correlated.

Table 3: Spatial autocorrelation measured by Moran's I for community weighted means and variances. _Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations. FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r SAC}
library(ape)
traits <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
traits <- c(paste0(traits, '.cwm'), paste0(traits, '.cwv'))
traits <- list(c(traits, 'FRic.fd', 'FEve.fd', 'FDiv.fd', 'FDis.fd', 'RaoQ.fd'), traits, traits)
weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
diag(weights) <- 0
SAC <- mapply(function(x,y){lapply(as.list(x@data[y]), Moran.I, weights)}, x =com, y = traits)
SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
SAC <- data.frame(SAC)
names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
SAC$Type <- ''
SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
kable(SAC, row.names = F)
```

```{r 6 lm & null models}
reg <- '~ Slope + Curvature + Wetness + SW + Canopy + BA'
# LM to simplify with loop or lapply
LM <- lapply(com, function(x) {with(x@data, list(LES.cwm = lm(formula(paste('LES.cwm', reg))), # CWM
                                                 logSLA.cwm = lm(formula(paste('log(SLA.cwm)', reg))), 
                                                 WES.cwm = lm(formula(paste('WES.cwm', reg))),
                                                 WD.cwm = lm(formula(paste('WD.cwm', reg))),
                                                 LDMC.cwm = lm(formula(paste('LDMC.cwm', reg))),
                                                 logLA.cwm = lm(formula(paste('log(LA.cwm)', reg))),
                                                 Thick.cwm = lm(formula(paste('Thick.cwm', reg))),
                                                 LES.cwv = lm(formula(paste('LES.cwv', reg))), # CWV
                                                 logSLA.cwv = lm(formula(paste('log(SLA.cwv)', reg))), 
                                                 WES.cwv = lm(formula(paste('WES.cwv', reg))),
                                                 WD.cwv = lm(formula(paste('WD.cwv', reg))),
                                                 LDMC.cwv = lm(formula(paste('LDMC.cwv', reg))),
                                                 logLA.cwv = lm(formula(paste('log(LA.cwv)', reg))),
                                                 Thick.cwv = lm(formula(paste('Thick.cwv', reg)))))})
LM$abundance <- c(LM$abundance, with(com$abundance@data, list(FRic.fd = lm(formula(paste('FRic.fd', reg))),
                                                              FEve.fd = lm(formula(paste('FEve.fd', reg))),
                                                              FDiv.fd = lm(formula(paste('FDiv.fd', reg))),
                                                              FDis.fd = lm(formula(paste('FDis.fd', reg))),
                                                              RaoQ.fd = lm(formula(paste('RaoQ.fd', reg))))))
LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
model <- null_model(com, Trees, PFT_sp, lapply(LM, function(y){lapply(y, function(x){x$call})}), SpCode, n = 999, verbose = F, time = T)
```

## Community weighted means

Table 4: Summary for linear models and null models for traits community weighted means. _Linear model results for each trait community weighted means (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA). Model deviance was compared to 999 null model deviances in which traits were randomized among species. See table 1 for abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r CWM}
kable(regression_table(LM, model, reg, select = '.cwm'))
```

### Abiotic factors

First we can notice that, whereas the leaf economic spectrum shows significantly non-random links with abiotic factors, most of weighted means associated to the wood economic spectrum on our principal component analysis (see figure 1) didn't differ from null models. A spatial structure due to local individual history and dispersion is responsible for the observed links between abiotic factors and weighted mean associated to the wood economic spectrum. Outside from the leaf economic spectrum (LES), topography doesn't localy afect Uppangala plot trees as an environmental filter.

Leaf economic spectrum (LES) and specific leaf area (SLA) is associated non-randomly with convex and dry surfaces, as the thickness for basal area community weighted mean. Trees will tend to increase their surface compare to their carbon invest ment in areas where the water will flow but not accumulate nor stay, and where the soil will tend to accumulate and be stabilized. In favourable water and soil conditions, trees will tend to invest more in leaf ressource capture at the expense of leaf defense, adopting the acquisitive [@sterck_functional_2011] our fast growing strategy [@reich_world-wide_2014].

Wood density (WD) in presence absence shows a non-random association with southwest steep slope and concave surfaces with a relatively important explanation level ($R^2=0.25$). Trees will increase their wood density in thin soil and water and soil departure conditions under potential monsoon winds *(We need to decide or not if regarding the plot condition we can accept or not the southwest orientation as being linked to monsoon winds)*. In water stress and mechanical hazard conditions, trees will compensate by increasing their water transport safety and mechanical resistance and resilience with an increased wood density [@chave_towards_2009].

### Biotic factors

Most of community weighted means don't show association with biotic factors, and most of the rare association are not significantly different from null models randomly generated. Still leaf area (LA) in basal area is associated with high trees. Voluminous individuals tend to increase their leaf surface in heigher canopy condition to compensate understory shadow. And wood density tends also to increase with tree height for rare species, suggesting that with higher canopy denser wood species will tend to grow in a community, as shade species.

## Community weighted variances

Table 5: Summary for linear models and null models for traits community weighted variances. _Linear model results for each trait community weighted variances (by abundance (AB), presence-absence (PA) and basal area (BA)) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA). Model deviance was compared to 999 null model deviances in which traits were randomized among species. See table 1 for abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r CWV}
kable(regression_table(LM, model, reg, select = '.cwv'))
```

Most of traits and tree strategy axes didn't show linear models for community weighted variances (CWV) significantyl differently affected by both abiotic and biotic environment compare to null models. Only the leaf economic spectrum (LES) and wood density (WD) showed linear models significantly different from null models for their community weighted variances (CWV) for all wieghts (abundance, presence-absence, and basal area). Leaf area community weighted variance by basal area showed also a significantly positive linear model with slope compare to null models.

Leaf economic spectra (LES) converge with better soil and water conditions in convex situations, and for abundant species oriented under monsoon wind LES diverge suggesting different adaptive strategies from species. Whereas wood density from abundant and rare species also diverged with the southwest orientation to monsoon winds, wood density from voluminous species converge with slope and diverge with the canopy height. Effectively under high canopy trees will be under a lot of shade situation, implying differing growth strategies and further more different wood densities. Leaf area also converge for voluminous species with slope.

## Functionnal diversity

Whereas functionnal richness (FRic) standardized by the *'global FRic'* that include all species is low and relatively stable (mean = 0.05, sd = 0.04);
functionnal eveness is high and relatively stable (mean = `r round(mean(com$abundance@data$FEve.fd), 2)`, sd = `r round(sd(com$abundance@data$FEve.fd), 2)`) in our communities as the functionnal diversity (mean = `r round(mean(com$abundance@data$FDiv.fd), 2)`, sd = `r round(sd(com$abundance@data$FDiv.fd), 2)`). Our communties are relatively poor in functionnal values but with an even and diverse trait distribution.

Table 6: Summary for linear models and null models for functionnal diversity indices (FD). _Linear model results for each community functionnal indices weighted by abundances after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA). Model deviance was compared to 999 null model deviances in which traits were randomized among species. FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r FD}
kable(regression_table(LM, model, reg, select = '.fd')[-2], row.names = F)
```

At the exception of the functionnal richness (FRic), all linear models for functionnal diversity indices didn't differ from null models suggesting that our bioctic and abitotic factors are not responsible for their distribution in our communities. Still functionnal richness increases with slope, suggesting as commmunity weighted variances (CWV) that our traits diverge in different adaptive strategies in steeper slopes.

## Models residuals

Table 6: Spatial autocorrelation measured by Moran's I for linear regression model residuals from community weighted means (CWM) and variances (CWV) and functionnal diversity indices (FD). _Spatial autocorrelation for residuals from each trait community weighted means and variances (by abundance, presence-absence and basal area) linear models and for functionnal diversity (FD) linear models, see table 1 for abbreviations. FRic, FEve, FDiv, FDis  and RaoQ stand respectively for functionnal richness, eveness, diversity dispersion and  Rao's quadratic entropy. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

```{r SAC residuals}
library(ape)
weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
diag(weights) <- 0
SAC <- lapply(LM, function(x){lapply(x, function(y){Moran.I(residuals(y), weights)})})
SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
SAC <- lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))})
SAC[[2]] <- c(SAC[[2]], rep('', 5)) ; SAC[[3]] <- c(SAC[[3]], rep('', 5))
SAC <- data.frame(SAC)
names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
SAC$Indice <- c(rep(c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA'), 2), c('FRic', 'FEve', 'FDiv', 'FDis', 'RaoQ'))
SAC$Type <- ''
SAC$Type[c(1,8,15)] <- c('CWM', 'CWV','FD')
SAC <- SAC[c('Type', 'Indice', 'Abundance', 'Presence-absence', 'Basal area')]
kable(SAC, row.names = F)
```

Model residuals spatial autocorrelation (SAC)  for our community weighted means (CWM) is null confirming that our models explain fully the spatial structure of our 5 traits and 2 strategy axes (compare to table 3). At the exception of thickness that still shows a significant spatial autocorrelation, suggesting that our model missed spatially structured factors in this case. For models residuals from community weighted variances, specific leaf area, wood economic spectrum and leaf dry matter content still show a spatial structure. This spatial structure in local trait variances linked to alpha diversity may be explained by local history and stochasticty associated with a null distribution of individuals not directly shaped by their environment. Finally we find back this spatial structure unexplained by model for functionnal diversity indices.

@@

```{r phylosignal, include=FALSE}
library(phylosignal)
library(phytools)
library(phylobase)

@@

# Phylosignal
tree <- results$Scenario.1
data <- PFT_sp
row.names(data) <- sub(' ', '_', row.names(data), fixed = T)
data <- data[tree$tip.label,c(2:5,7,9:10)]
p4d <- phylo4d(tree, data)
barplot.phylo4d(p4d, tree.type = "phylo", tree.ladderize = TRUE)
# phylosim <- phyloSim(tree = tre, method = "all", nsim = 100, reps = 99)
# plot.phylosim(phylosim, what = "pval", stacked.methods = TRUE)
phyCor <- lapply(as.list(names(data)), function(x){phyloCorrelogram(p4d, trait = as.character(x))})
# par(mfrow = c(2,4))
# for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
lipa <- lipaMoran(p4d)
# barplot.phylo4d(p4d, bar.col=(lipa$p.value < 0.05) + 1, center = F, scale = F)
```

>>>>

---
title: 'Uppangala tree functionnal traits: RLQ and Fourth cornr analysis'
author: "Sylvain Schmitt"
date: '`r Sys.Date()`'
output: pdf_document
bibliography: refs.bib
---

@@

## RLQ analysis

Following @dray_testing_2008, we conducting both RLQ and fourthcorner analysis to test the effect of environment on our traits. Model used to compare our results with a null model in both RLQ Monte-Carlo test and in the fourth corner analysis is the model 6 proposed by @ter_braak_improved_2012.

```{r RLQ}
library(ade4)

# Matrices
Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0
Qm <- PFT_sp[c('SLA', 'LDMC', 'WD', 'Thick', 'LA')]
row.names(Qm) <- PFT_sp$Sp_Code
Qm <- Qm[-which(is.na(Qm$WD)),]
Qm <- Qm[names(Lm),]
Rm <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(Rm) <- com@data$id

# Consistency
# all(row.names(Lm) == row.names(Rm))
# all(row.names(Lm) %in% row.names(Rm))
Lm <- Lm[row.names(Rm),]
# all(names(Lm) == row.names(Qm))
Qm <- Qm[names(Lm),]
Qm <- Qm[-which(is.na(Qm$WD)),]
Lm <- Lm[row.names(Qm)]
Lm <- Lm / rowSums(Lm)

# RLQ
L <- dudi.coa(Lm, scannf = F)
R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
rlq <- rlq(R, L, Q, scannf = F)

# Rersults
par(mfrow = c(1,2))
s.arrow(rlq$l1)
s.arrow(rlq$c1)
# plot(rlq)
```

Figure 3: Environmental variables (left) and traits (right) ordination with RLQ analysis results for L matrix in abundances.

```{r randtest}
rd <- ade4::randtest(rlq, modeltype = 6)
plot(rd)
```

Figure 4: RLQ Monte-Carlo test results for model 6 proposed by @ter_braak_improved_2012.

## Fourth corner analysis

```{r 4thQ AB}
library(ade4)
# FQR
fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")

# Results
par(mfrow = c(1,2))
plot(fqr)
plot(fqr.adj)
# print(fqr.adj)
```

Figure 5: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species abundances in communities with non adjusted p-values (left) and adjusted p-values (right).

```{r 4thQ BIN}
library(ade4)

# Matrices
Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0

# Consistency
Lm <- Lm[row.names(Rm),]
Lm <- Lm[row.names(Qm)]
Lm[Lm > 0] <-  1 #Presence-absence
Lm <- Lm / rowSums(Lm)

# RLQ
L <- dudi.coa(Lm, scannf = F)
R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
rlq <- rlq(R, L, Q, scannf = F)
fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")

# Results
par(mfrow = c(1,2))
plot(fqr)
plot(fqr.adj)
# print(fqr.adj)
```

Figure 6: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species presence absence in communities with non adjusted p-values (left) and adjusted p-values (right).

```{r 4thQ BA}
library(ade4)

# Matrices
Lm <- data.frame(tapply(Trees$`2013_girth`/2*pi, list(Trees$com, Trees$SpCode), sum)) #BA
Lm[is.na(Lm)] <- 0

# Consistency
Lm <- Lm[row.names(Rm),]
Qm <- Qm[names(Lm),]
Qm <- Qm[-which(is.na(Qm$WD)),]
Lm <- Lm[row.names(Qm)]
Lm <- Lm / rowSums(Lm)

# RLQ
L <- dudi.coa(Lm, scannf = F)
R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
Q <- dudi.pca(Qm, scannf = F, row.w = L$cw)
rlq <- rlq(R, L, Q, scannf = F)
fqr <- fourthcorner(Rm, Lm, Qm, modeltype = 6, p.adjust.method.G = "none", p.adjust.method.D = "none")
fqr.adj <- p.adjust.4thcorner(fqr, p.adjust.method.G = "fdr", p.adjust.method.D = "fdr")

# Results
par(mfrow = c(1,2))
plot(fqr)
plot(fqr.adj)
# print(fqr.adj)
```

Figure 7: Fourth corner analysis with model 6 proposed by @ter_braak_improved_2012 with matrix L based on species basal area in communities with non adjusted p-values (left) and adjusted p-values (right).

@@

```{r phylosignal}
library(phylosignal)
library(phytools)
library(phylobase)

@@

# Phylosignal
tree <- results$Scenario.1
data <- PFT_sp
row.names(data) <- sub(' ', '_', row.names(data), fixed = T)
data <- data[tree$tip.label,c(2:5,7,9:10)]
p4d <- phylo4d(tree, data)
# barplot.phylo4d(p4d, tree.type = "phylo", tree.ladderize = TRUE)
# phylosim <- phyloSim(tree = tre, method = "all", nsim = 100, reps = 99)
# plot.phylosim(phylosim, what = "pval", stacked.methods = TRUE)
phyCor <- lapply(as.list(names(data)), function(x){phyloCorrelogram(p4d, trait = as.character(x))})
par(mfrow = c(2,4))
for(i in 1:length(names(data))) plot(phyCor[[i]], main = names(data)[i])
lipa <- lipaMoran(p4d)
# barplot.phylo4d(p4d, bar.col=(lipa$p.value < 0.05) + 1, center = F, scale = F)
```

Once the phylogenetic structure revealed we wanted to test it in our community with the Bst and PIst metric proposed by @hardy_characterizing_2007 with the null model 1s suggested by [cite] to see if we were under phylogenetic clustering or overdispersion.

```{r phylo structure}
library(spacodiR)

# Phylogeny renaming
tree$tip.label <- gsub('_', ' ', tree$tip.label, fixed = T)
tree$tip.label <- as.character(SpCode[tree$tip.label])

# Matrices
Lm <- data.frame(tapply(rep(1, length(Trees[,1])), list(Trees$com, Trees$SpCode), sum))
Lm[is.na(Lm)] <- 0
Rm <- com@data[c('Slope', 'Curvature', 'Wetness', 'SW', 'Canopy', 'BA')]
row.names(Rm) <- com@data$id

# Consistency
# all(row.names(Lm) == row.names(Rm))
# all(row.names(Lm) %in% row.names(Rm))
Lm <- Lm[row.names(Rm),]
# all(names(Lm) == tree$tip.label)
Lm <- Lm[tree$tip.label]

# Phylogeny structure
spafun <- spacodi.calc(sp.plot = t(Lm), sp.traits = Qm)
spaphy <- spacodi.calc(sp.plot = t(Lm), phy = tree)
permut <- lapply(as.list(names(spaphy[1:4])), function(x){spacodi.by.nodes(t(Lm), tree, sp.parm = x, n.rep = 100, method = "1s")})
names(permut) <- names(spaphy[1:4])
par(mfrow=c(2,1))
spacodi.permutplot(permut$Bst)
spacodi.permutplot(permut$PIst)
```

The phylogenetic structure among our communities was then evaluated by an RLQ analysis with cophenetic matrix of our phylogenetic tree distances as the Q matrix reduced by a principal coordinate analysis.

```{r phylo RLQ}
library(ade4)
library(adephylo)

# RLQ
L <- dudi.coa(Lm, scannf = F)
R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
Qd <- distTips(tree)
Q <- dudi.pco(Qd, scannf = F, row.w = L$cw)
rlq <- rlq(R, L, Q, scannf = F)

# Random RLQs
tree$node.label <- NULL
ntrees <- rep(list(tree), 999)
ntrees <- lapply(ntrees, resamp.phy)
nQd <- lapply(ntrees, distTips)
nrlq <- lapply(ntrees, function(x){
  Lm <- Lm[x$tip.label]
  L <- dudi.coa(Lm, scannf = F)
  R <- dudi.pca(Rm, scannf = F, row.w = L$lw)
  Q <- dudi.pco(Qd, scannf = F, row.w = L$cw)
  rlq(R, L, Q, scannf = F)})

# Randtest
pval <- rank(c(sum(rlq$eig), unlist(lapply(nrlq, function(x){sum(x$eig)}))))[1]/1000
hist(unlist(lapply(nrlq, function(x){sum(x$eig)})), main = paste('pval =', round(pval, 3)))
lines(c(sum(rlq$eig), sum(rlq$eig)), c(0,500), col = 'red', lwd = 2)
```

Figure 8: RLQ test by model nul results with phylogenetic tree distances as the Q matrix for model 1s proposed by [cite].

### 2016-07-01 inst/doc/draft.Rmd
v0.1.9016 RLQ.Rmd + cosaspect fix

---
title: 'Uppangala tree functionnal traits: Draft'
author: "Sylvain Schmitt"
date: '`r Sys.Date()`'
output: pdf_document
bibliography: refs.bib
---

@@

The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village. The permanent plot was established in 1990 by the French Institute of Pondicherry and described by Pascal and Pelissier [@pascal_structure_1996]. The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergree forest type [@pascal_wet_1988]. Climate and geology of the 10-ha plot can be considered as homogeneous. General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].

@@

We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates. Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals. Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot. For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013. Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy. Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling.

- Unsampled species beanch wood 
- Leaves manipulation conditions.

Among leaves, 5 were weighted fresh, scanned [cite scan], and their thickness were measured with a microgauge [cite]. Those leaves were kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying processed was stopped after dry leaf weights were stable in time. Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias. Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues. Branch wood samples were then sink in water to evaluate their volume by water displacement. Samples were after dried for 48 hours at 80 degrees celsius and weighted.

>>>>

---
title: 'Uppangala tree functionnal: Article draft'
author: Sylvain Schmitt sylvain.schmitt@agroparistech.fr
output: pdf_document
bibliography: refs.bib
---

@@

The study was carried out in the Kadamakal Reserve Forest (Kodagu District, Karnataka State, India), in the 10-ha "Uppangala plot" named after the nearest village. The permanent plot was established in 1990 by the French Institute of Pondicherry and described by @pascal_structure_1996. The Uppangala plot vegetation was assigned to low elevation *Dipterocarpus indicus - Kingiodendron pinnatum - Humboldtia brunonis* dense moist evergree forest type [@pascal_wet_1988]. Climate and geology of the 10-ha plot can be considered as homogeneous. General topagraphy of the site shows a regular east-west alternation between thalwegs and ridges separated by relatively steep slopes, so that topography can be assumed as the main local abiotic factor which varies inside the plot [@gimaret-carpentier_sampling_1998].

@@

We collected trait information on 92 over 117 (79%) species focusing in priority on most abundant species with 5 replicates and we completed with less abundant and rare species with one to five replicates. Most abundant species were defined as the 32 species respresenting more than 80% of Uppangala tree individuals. Thus we generated data for 98.8% of the tree individuals present in the Uppangala plot. For each individual we collected 6 mature shadow dry leaves and one branch sample with a diameter close to 1cm following recommendation from @perez-harguindeguy_new_2013. Individuals and leaves were selected regarding their light condition, development stage and sanitary state for consistancy. Dawkins crown exposure index [@dawkins_management_1958] was registered to check the validity of the sampling. Leaves and branch samples were kept in humidified ziplock filled with enriched CO2 air inside a black bag untill their measurement in the following 24 hours. For some individuals we were unable to collect a correct branch sample (58 over 267 individuals representing 12 species over 91).

Leaf fresh weights, area and thickness were measured, using a scanner [cite] and a microgauge [cite]. Leaves were first kept in herbarium and then dried in an oven for at least 72 hours at 60 degrees celsisus, the drying process was stopped after dry leaf weights were stable in time. Once dried the leaves were weighted directly after been removed from the oven to avoid humidity bias. Branch samples bark and pit were removed in order to keep only branch wood with no soft tissues. Branch wood samples were then sink in water to evaluate their volume by water displacement. Samples were after dried for 48 hours at 80 degrees celsius and weighted.

- Removed petiole, petiolule and rachis
